Poincar\'e--Friedrichs constants for the $L^{p}$ de Rham complex over domains with shellable triangulations

### 2024-06-28 main.tex
rename variables; Gradient section

\title{Poincar\'e--Friedrichs constants over domains with shellable triangulations}
\author{
    Th\'eophile Chaumont-Frelet\thanks{Affiliation of the First Author, \texttt{email1@example.com}} \and
    Martin Werner Licht\thanks{Affiliation of the Second Author, \texttt{email2@example.com}} \and
    Martin Vohral\'ik\thanks{Affiliation of the Third Author, \texttt{email3@example.com}}
}

@@

Finding a potential operator, whether linear or non-linear, and bounding the Poincar\'e--Friedrichs constant is generally not an easy undertaking. 
% We accomplish our results for a special class of triangulations, so-called shellable triangulations.
We accomplish this undertaking in the special case where $\Omega$ admits a so-called \emph{shellable} triangulation. 
Even though only contractible domains can ever admit a shellable triangulation, 
having computable upper bounds for such domains is an important stepping stone towards the more general case. 
For example, local patches around a simplex within a larger 2D or 3D triangulation are shellable.
We express all our main results in the calculus of differential forms, also known as exterior calculus, 
and treat the curl and divergence operator as its particular instances. 
\\

@@

\section{Constructive estimate of Poincar\'e constants}\label{section:poincare}

@@

Let us assume that $\Omega \subseteq \bbR^{n}$ is an open set. 
Given any $p \in [1,\infty]$, we let $L^p(\Omega)$ denote the Lebesgue space over $\Omega$ with integrability exponent $p$, and we write $\bfL^p(\Omega) := L^p(\Omega)^{n}$ for the corresponding Lebesgue space of vector fields. 
We also write $W^{1,p}(\Omega)$ for the first-order Sobolev space over $\Omega$ with integrability exponent $p$.

We say that a domain $\Omega \subseteq \bbR^{n}$ satisfies the \emph{Poincar\'e--Friedrichs inequality} with exponent $p \in [1,\infty]$
if there exists a constant $C_{{\rm PF},\Omega,p} \geq 0$ such that the following holds:
for every vector field $\bff \in \nabla W^{1,p}(\Omega)$ there exists $u \in W^{1,p}(\Omega)$
such that $\nabla u = \bff$ and 
\begin{align}\label{math:inequalitypf:gradient}
    \| u \|_{L^{p}(\Omega)}
    \leq 
    C_{{\rm PF},\Omega,p} 
    \| \bff \|_{L^{p}(\Omega)}
    .
\end{align}
Equivalently, for all $u \in W^{1,p}(\Omega)$ we have 
\begin{align}
    \min_{ c \in \bbR } \| u - c \|_{L^{p}(\Omega)}
    \leq 
    C_{{\rm PF},\Omega,p} 
    \| \nabla u \|_{L^{p}(\Omega)}
    .
\end{align}
We call $C_{{\rm PF},\Omega,p}$ the \emph{Poincar\'e--Friedrichs constant} with exponent $p$.

We wish to clarify the relationship between the Poincar\'e--Friedrichs inequality, in the sense introduced above, with other inequalities that are known as Poincar\'e inequality (or also Poincar\'e--Wirtinger or Friedrichs inequality) in the literature. Given $p \in [1,\infty]$ and an open set $\Omega \subseteq \bbR^{n}$ of finite measure, 
we say that $\Omega$ satisfies the Poincar\'e inequality with exponent $p$ 
if there exists $C_{{\varnothing},\Omega,p} \geq 0$ such that 
for all $u \in W^{1,p}(\Omega)$
\begin{align}
    \| u - u_{\Omega} \|_{L^{p}(\Omega)}
    \leq 
    C_{\varnothing,\Omega,p} 
    \| \nabla u \|_{L^{p}(\Omega)}
    ,
\end{align}
where $u_{\Omega}$ is the \emph{average} of $u$ over $\Omega$, that is,
\begin{align*}
    u_{\Omega} := |\Omega|^{-1} \int_{\Omega} u.
\end{align*}
Clearly, this Poincar\'e inequality implies the Poincar\'e--Friedrichs inequality and we have $C_{{\rm{PF}},\Omega,p} \leq C_{\varnothing,\Omega,p}$. 
Towards a converse inequality, 
let us first observe that the average of any $v \in W^{1,p}(\Omega)$ with $p < \infty$ satisfies the bound 
\begin{align*}
    \| v_\Omega \|_{L^{p}(\Omega)}^{p}
    = 
    \int_{\Omega} \left( |\Omega|^{-1} \int_{\Omega} |v(x)| \right)^{p}
    \leq 
    \int_{\Omega} \left( |\Omega|^{-1} \int_{\Omega} |v(x)|^{p} \right)
    = 
    \| v \|_{L^{p}(\Omega)}^{p}
    .
\end{align*}
Here, we have used Jensen's inequality. 
In the case $p = \infty$, any $v \in L^{\infty}(\Omega)$ satisfies $\| v_\Omega \|_{L^{\infty}(\Omega)} \leq \| v \|_{L^{\infty}(\Omega)}$. 
It follows that 
\begin{align*}
    \| v - v_\Omega \|_{L^{p}(\Omega)} 
    \leq
    2
    \| v \|_{L^{p}(\Omega)},
    \quad 
    v \in L^{p}(\Omega)
    .
\end{align*}
% \begin{align*}
%     \color{Emerald}
%     \| u - u_\Omega \|_{L^{p}(\Omega)}^{p}
%     \leq 
%     2^{p-1} 
%     \| u \|_{L^{p}(\Omega)}^{p}
%     +
%     2^{p-1} 
%     \| u_\Omega \|_{L^{p}(\Omega)}^{p}
%     \leq 
%     2^{p} 
%     \| u \|_{L^{p}(\Omega)}^{p}
%     .
% \end{align*}
Hence the Poincar\'e--Friedrichs inequality implies the Poincar\'e inequality with $C_{\varnothing,\Omega,p} \leq 2 C_{{\rm PF},\Omega,p}$.

In the special case $p=2$, we always have $\| u - u_\Omega \|_{L^{2}(\Omega)} \leq \| u \|_{L^{2}(\Omega)}$ for any $v \in L^{2}(\Omega)$,
and therefore even $C_{\varnothing,\Omega,2} = C_{{\rm PF},\Omega,2}$. 
This follows from the orthogonality constant functions to average-free functions in $L^2(\Omega)$,
and it also follows from the projection estimate (see, e.g.,~\cite{xu2003some}).
%\footnote{\color{red}Is there a better name for this than Xu-Zikatanov estimate?}
Stern~\cite{stern2015banach} provided a generalization of the projection estimate that implies improved estimate for all $L^p$ spaces with $1 \leq p \leq \infty$:
since taking the average is a the projection onto the constants functions with unit norm, it now follows that 
\begin{align*}
    \| v - v_\Omega \|_{L^{p}(\Omega)}
    \leq 
    \min\left( 2, 2^{|2/p-1|} \right)
    \| v \|_{L^{p}(\Omega)}
    = 
    2^{|2/p-1|} 
    \| v \|_{L^{p}(\Omega)}
    ,
    \quad 
    v \in L^p(\Omega)
    .
\end{align*}
We have $1 \leq 2^{|2/p-1|} \leq 2$ for $1 \leq p \leq \infty$.
The upper bound is only attained in the limit cases $p = 1$ and $p = \infty$, and the lower bound is only attained if $p = 2$, which reproduces the improved estimate for the Hilbert space case.
The main consequence for us is the estimate 
$C_{\varnothing,\Omega,p} \leq 2^{|2/p-1|} C_{{\rm PF},\Omega,p}$.
We conclude that our notion of Poincar\'e--Friedrichs constant is equivalent to the notion of Poincar\'e constant in the literature, up to a numerical factor that only depends on $1 \leq p \leq \infty$.

There is yet another characterization of the Poincar\'e--Friedrichs inequality that we like to point out. 
We define the potential 
\begin{align*}
    \Phi( \bff ) := \argmin\limits_{ \substack{ v \in W^{1,p}(\Omega) \\ \nabla v = \bff } } \| v \|_{L^{p}(\Omega)},
    \quad 
    \bff \in \nabla W^{1,p}(\Omega)
    .
\end{align*}
Clearly, 
\begin{align*}
    \| \Phi(\bff) \|_{L^p(\Omega)} \leq C_{{\rm PF},\Omega,p} \| \bff \|_{L^p(\Omega)},
\end{align*}
and this inequality is sharp by definition of $\Phi$ provided that $C_{{\rm PF},\Omega,p}$ is smallest possible constant in \eqref{math:inequalitypf:gradient}. 
However, $\Phi : \nabla W^{1,p}(\Omega) \rightarrow L^p(\Omega)$ is generally a nonlinear operator for $p \neq 2$. 
Consequently, any \textit{linear} potential operator $\overline\Phi : \nabla W^{1,p}(\Omega) \rightarrow L^p(\Omega)$
satisfying $\nabla \overline\Phi( \bff ) = \bff$ for any $\bff \in \nabla W^{1,p}(\Omega)$ must have an operator norm that obeys the lower bound 
\begin{align*}
    C_{{\rm PF},\Omega,p} 
    \leq 
    \argmin\limits_{ v \in \nabla W^{1,p}(\Omega) \setminus \bbR } 
    \frac{ \| \overline\Phi(\nabla v) \|_{L^p(\Omega)} }{ \| \nabla v \|_{L^p(\Omega)} }
    .
\end{align*}
A natural choice is the linear operator $\Phi_{\varnothing} : \nabla W^{1,p}(\Omega) \rightarrow L^p(\Omega)$ that satisfies 
\begin{align*}
    \Phi_{\varnothing}( \nabla u ) = u - u_{\Omega},
    \quad 
    u \in W^{1,p}(\Omega)
    .
\end{align*}
Its operator norm is just the smallest Poincar\'e constant $C_{\varnothing,\Omega,p}$.

In the light of these musings, we clarify that our objective is the construction of linear potential operators for the gradient and later for the curl and divergence operators,
which is more specific than just proving an upper bound for the Poincar\'e--Friedrichs constant. 
Any Poincar\'e--Friedrichs inequality~\eqref{math:inequalitypf:gradient} is comparable to an upper bound for the generalized (possibly nonlinear) inverse of the gradient operator $\nabla : W^{1,p}(\Omega) \rightarrow L^{p}(\Omega)^{n}$. 
The potential reconstruction is the natural approach towards the topic when addressing the curl and divergence operators.

\begin{remark}
    Let us further remark why the above notion of Poincar\'e--Friedrichs inequality suits our discussion better than the Poincar\'e inequality. 
    We want to generalize the discussion to the curl and divergence operators. 
    The kernel of the gradient is the one-dimensional space of constant functions, and is thus complemented in the Lebesgue spaces with a canonical choice of projection. 
    By contrast, the curl and divergence operators have infinite-dimensional kernels. 
    Hence, it is not clear a priori whether these kernels are complemented subspaces and admit a projection onto them, not to mention a canonical projection. 
    % The kernel of those operators is infinite-dimensional, whereas the gradient only has a one-dimensional kernel. 
    % When discussing Banach spaces, there is no canonical orthogonal projection on the kernel, unlike in the Hilbert space case, 
    % which is why the Poincar\'e--Friedrichs inequality seems to be the more natural approach.
\end{remark}

@@

\subsection{Poincar\'e inequalities over convex domains}
% See the collection at Lemma 3.24 : https://hal.science/hal-03226049/file/volI_hal.pdf

We now collect examples for Poincar\'e--Friedrichs inequalities for important special cases; see also~\cite[Lemma~3.24]{ern2021finite} for a short overview. 
Over convex bounded domains, we have the Poincar\'e inequalities~\cite{bebendorf2003note,acosta2004optimal} 
\begin{align}\label{math:acostaduran}
    \| u - u_{\Omega} \|_{L^{1}(\Omega)}
    \leq 
    \frac{\diam(\Omega)}{2}
    \| \nabla u \|_{L^{1}(\Omega)}
    ,
    \quad 
    u \in W^{1,1}(\Omega)
    ,
    \\
    \| u - u_{\Omega} \|_{L^{2}(\Omega)} \label{math:bebendorf}
    \leq 
    \frac{\diam(\Omega)}{\pi}
    \| \nabla u \|_{L^{2}(\Omega)}
    ,
    \quad 
    u \in W^{1,2}(\Omega)
    .
\end{align}
These two estimates are the best possible in the cases $p=1$ and $p=2$, respectively, in terms of the diameter alone. 
Upper bounds for the Poincar\'e constant over convex domains with $1 < p < \infty$ are known in the literature~\cite{chua2006estimates}:
\begin{align}\label{math:chuawheeden}
    \| u - u_{\Omega} \|_{L^{p}(\Omega)}
    \leq 
    C_{{\rm CW},p}
    \diam(\Omega)
    \| \nabla u \|_{L^{p}(\Omega)}
    ,
    \quad 
    u \in W^{1,p}(\Omega)
    ,
\end{align}
where 
\begin{align}
    C_{{\rm CW},p} 
    := 
    \sup\limits_{ v \in C^\infty([0,1]) \setminus \bbR } 
    \frac{ 
        \| v - v_{[0,1]} \|_{L^{p}([0,1])} 
    }{ 
        \| \nabla v \|_{L^{p}([0,1])} 
    }
    \leq 
    \sqrt[p]{p} 2^{1-\frac 1 p}
    =
    2
    \left( \frac p 2 \right)^{\frac 1 p}
    .
\end{align}
The literature knows optimal Poincar\'e--Friedrichs constants over convex domains. 
According to~\cite{ferone2012remark,esposito2013poincare}, when $1 < p < \infty$, one can show that 
\begin{align}\label{math:esposito:pf}
    \min\limits_{ c \in \bbR }
    \| u - c \|_{L^{p}(\Omega)}
    \leq 
    C_{\rm{EFNT},p}
    \diam(\Omega)
    \| \nabla u \|_{L^{p}(\Omega)}
    ,
    \quad 
    u \in W^{1,p}(\Omega)
    ,
\end{align}
where $C_{\rm{EFNT},p}$ is the best possible constant that only depends on $p$ and equals 
\begin{align}
    C_{\rm{EFNT},p}
    :=
    \frac{ p \sin(\pi/p) }{ 2\pi \sqrt[p]{p-1} }
    .
\end{align}
Note that the last inequalities imply, again when $1 < p < \infty$, the Poincar\'e inequalities
\begin{align}\label{math:esposito:poincare}
    \| u - u_{\Omega} \|_{L^{p}(\Omega)}
    \leq 
    2^{|1-\frac 2 p|}C_{\rm{EFNT},p}
    \diam(\Omega)
    \| \nabla u \|_{L^{p}(\Omega)}
    ,
    \quad 
    u \in W^{1,p}(\Omega)
    .
\end{align}
The main feature in proving these results is that the best constant in the Poincar\'e inequality is already the constant in the $1$-dimensional case. 
Finally, since convex domains are Lipschitz domains: 
\begin{align}\label{math:lipschitzestimate}
    \min\limits_{ c \in \bbR }
    \| u - c \|_{L^{\infty}(\Omega)}
    \leq 
    \diam(\Omega)
    \| \nabla u \|_{L^{\infty}(\Omega)}
    ,
    \quad 
    u \in W^{1,\infty}(\Omega)
    .
\end{align}

\begin{remark}
    Let us compare the different estimates for the Poincar\'e inequality.
    In the case $2 \leq p$, 
    \begin{align*}
        2^{1-\frac 2 p}
        C_{\rm{EFNT},p}
        =
        \frac{ 2^{1-\frac 2 p} }{2}
        \frac{ \sin(\pi/p) }{ \pi/p }
        \frac{ 1 }{ \sqrt[p]{p-1} }
        =
        4^{-\frac 1 p}
        \leq
        C_{{\rm CW},p} 
        .
    \end{align*}
    In the case $2 \geq p$, 
    \begin{align*}
        2^{\frac 2 p - 1}
        C_{\rm{EFNT},p}
        =
        \frac{ 2^{\frac 2 p - 1} }{2}
        \frac{ \sin(\pi/p) }{ \pi/p }
        \frac{ 1 }{ \sqrt[p]{p-1} }
        \leq
        \frac{ 2^{\frac 2 p-1} }{2}
        =
        2^{\frac 2 p - 2}
        =
        4^{\frac 1 p - 1}
        \leq
        C_{{\rm CW},p} 
%         % 
%         .
%         4^{\frac 1 p}
%         \frac{ p \sin(\pi/p) }{ 2\pi \sqrt[p]{p-1} \cdot 2 }
%         =
%         4^{\frac 1 p}
%         \frac{ p \sin(\pi/p) }{ 4\pi \sqrt[p]{p-1} }
%         \leq 
%         \frac 1 2
        .
    \end{align*}
    It follows that \eqref{math:esposito:poincare} is tighter estimate than \eqref{math:chuawheeden} for $1 < p < \infty$.
    In the limit as $p$ goes to $1$, \eqref{math:esposito:poincare} recovers \eqref{math:acostaduran}. 
    In the limit as $p$ goes to infinity, \eqref{math:esposito:poincare} implies the Poincar\'e inequality 
    \begin{align*}
        \| u - u_{\Omega} \|_{L^{\infty}(\Omega)}
        \leq 
        \diam(\Omega)
        \| \nabla u \|_{L^{\infty}(\Omega)}
        ,
        \quad
        u \in W^{1,\infty}(\Omega)
        %
    \end{align*}
    because the $L^{\infty}$-norm is the limit of the $L^{p}$-norm. 
    Of course, this Poincar\'e inequality implies a Poincar\'e--Friedrichs inequality with the same constant. 
    Notably, it is much less clear whether \eqref{math:esposito:pf} implies a limit case for Lipschitz functions. 
%     Taking the \emph{informal} limit also suggest the inequality 
%     \begin{align*}
%         \min\limits_{ c \in \bbR }
%         \| u - c \|_{L^{\infty}(\Omega)}
%         \leq 
%         \frac{d}{2}
%         \| \nabla u \|_{L^{\infty}(\Omega)}
%         ,
%         \quad
%         u \in W^{1,\infty}(\Omega)
%         ,
%     \end{align*}
%     however, we generally cannot exchange the minimum and the limit, so that is strictly informal. 
\end{remark}

>>>>

\title{Poincar\'e--Friedrichs constants for the $L^{p}$ de Rham complex over domains with shellable triangulations}
\author{
    Th\'eophile Chaumont-Frelet\thanks{Affiliation of the First Author, \texttt{email1@example.com}} \and
    Martin Werner Licht\thanks{Affiliation of the Second Author, \texttt{email2@example.com}} \and
    Martin Vohral\'ik\thanks{Affiliation of the Third Author, \texttt{email3@example.com}}
}

@@

Finding a potential operator, whether linear or non-linear, and bounding the Poincar\'e--Friedrichs constant is generally not an easy undertaking. 
The majority of estimates in the literature address only gradient potentials. 
% We accomplish our results for a special class of triangulations, so-called shellable triangulations.
We accomplish this undertaking in the special case where $\Omega$ admits a so-called \emph{shellable} triangulation. 
Even though only contractible domains can ever admit a shellable triangulation, 
having computable upper bounds for such domains is an important stepping stone towards the more general case. 
For example, local patches around a simplex within a larger 2D or 3D triangulation are shellable.
We express all our main results in the calculus of differential forms, also known as exterior calculus, 
and treat the curl and divergence operator as its particular instances. 
\\

@@

\section{Constructive estimate of Poincar\'e--Friedrichs constants}\label{section:poincare}

@@

Let us assume that $\Omega \subseteq \bbR^{n}$ is an open connected set. 
Given any $p \in [1,\infty]$, we let $L^p(\Omega)$ denote the Lebesgue space over $\Omega$ with integrability exponent $p$, and we write $\bfL^p(\Omega) := L^p(\Omega)^{n}$ for the corresponding Lebesgue space of vector fields. 
We also write $W^{1,p}(\Omega)$ for the first-order Sobolev space over $\Omega$ with integrability exponent $p$.

We say that a domain $\Omega \subseteq \bbR^{n}$ satisfies the \emph{Poincar\'e--Friedrichs inequality} with exponent $p \in [1,\infty]$
if there exists a constant $C_{{\rm PF},\Omega,p} \geq 0$ such that the following holds:
for every vector field $\bff \in \nabla W^{1,p}(\Omega)$ there exists $u \in W^{1,p}(\Omega)$
such that $\nabla u = \bff$ and 
\begin{align}\label{math:inequalitypf:gradient}
    \| u \|_{L^{p}(\Omega)}
    \leq 
    C_{{\rm PF},\Omega,p} 
    \| \bff \|_{L^{p}(\Omega)}
    .
\end{align}
Under the condition that $\Omega$ is connected, this is equivalent to  
\begin{align}
    \min_{ c \in \bbR } \| u - c \|_{L^{p}(\Omega)}
    \leq 
    C_{{\rm PF},\Omega,p} 
    \| \nabla u \|_{L^{p}(\Omega)},
    \quad 
    u \in W^{1,p}(\Omega)
    .
\end{align}
We call $C_{{\rm PF},\Omega,p}$ the \emph{Poincar\'e--Friedrichs constant} with exponent $p$. 
\\

We wish to clarify the relationship between the Poincar\'e--Friedrichs inequality, in the sense introduced above, with other inequalities that are known as Poincar\'e inequality (or also Poincar\'e--Wirtinger or Friedrichs inequality) in the literature. Given $p \in [1,\infty]$ and an open set $\Omega \subseteq \bbR^{n}$ of finite measure, 
we say that $\Omega$ satisfies the Poincar\'e inequality with exponent $p$ 
if there exists $C_{{\varnothing},\Omega,p} \geq 0$ such that 
for all $u \in W^{1,p}(\Omega)$
\begin{align}
    \| u - u_{\Omega} \|_{L^{p}(\Omega)}
    \leq 
    C_{\varnothing,\Omega,p} 
    \| \nabla u \|_{L^{p}(\Omega)}
    ,
\end{align}
where $u_{\Omega}$ is the \emph{average} of $u$ over $\Omega$, that is,
\begin{align*}
    u_{\Omega} := |\Omega|^{-1} \int_{\Omega} u.
\end{align*}
Clearly, this Poincar\'e inequality implies the Poincar\'e--Friedrichs inequality and we have $C_{{\rm{PF}},\Omega,p} \leq C_{\varnothing,\Omega,p}$. 
Towards a converse inequality, 
let us first observe that the average of any $u \in W^{1,p}(\Omega)$ with $p < \infty$ satisfies the bound 
\begin{align*}
    \| u_\Omega \|_{L^{p}(\Omega)}^{p}
    = 
    \int_{\Omega} \left( |\Omega|^{-1} \int_{\Omega} |u(x)| \right)^{p}
    \leq 
    \int_{\Omega} \left( |\Omega|^{-1} \int_{\Omega} |u(x)|^{p} \right)
    = 
    \| u \|_{L^{p}(\Omega)}^{p}
    .
\end{align*}
Here, we have used Jensen's inequality. 
In the case $p = \infty$, any $u \in L^{\infty}(\Omega)$ satisfies $\| v_\Omega \|_{L^{\infty}(\Omega)} \leq \| u \|_{L^{\infty}(\Omega)}$. 
It follows that 
\begin{align*}
    \| u - u_\Omega \|_{L^{p}(\Omega)} 
    \leq
    2
    \| u \|_{L^{p}(\Omega)},
    \quad 
    u \in L^{p}(\Omega)
    .
\end{align*}
% \begin{align*}
%     \color{Emerald}
%     \| u - u_\Omega \|_{L^{p}(\Omega)}^{p}
%     \leq 
%     2^{p-1} 
%     \| u \|_{L^{p}(\Omega)}^{p}
%     +
%     2^{p-1} 
%     \| u_\Omega \|_{L^{p}(\Omega)}^{p}
%     \leq 
%     2^{p} 
%     \| u \|_{L^{p}(\Omega)}^{p}
%     .
% \end{align*}
Hence the Poincar\'e--Friedrichs inequality implies the Poincar\'e inequality with 
\begin{align*}
    C_{\varnothing,\Omega,p} \leq 2 C_{{\rm PF},\Omega,p}.
\end{align*}
In the special case $p=2$, this improves to $\| u - u_\Omega \|_{L^{2}(\Omega)} \leq \| u \|_{L^{2}(\Omega)}$ for any $u \in L^{2}(\Omega)$,
and therefore even $C_{\varnothing,\Omega,2} = C_{{\rm PF},\Omega,2}$. 
This improvement follows from the orthogonality of constant functions to average-free functions in $L^2(\Omega)$,
and it also follows from the projection estimate (see, e.g.,~\cite{xu2003some}).
Stern~\cite{stern2015banach} provides a generalized projection estimate that implies improved estimate for all $L^p$ spaces with $1 \leq p \leq \infty$:
since taking the average is a the projection onto the constants functions with unit norm, it now follows that 
\begin{align*}
    \| u - u_\Omega \|_{L^{p}(\Omega)}
    \leq 
    \min\left( 2, 2^{|2/p-1|} \right)
    \| u \|_{L^{p}(\Omega)}
    = 
    2^{|2/p-1|} 
    \| u \|_{L^{p}(\Omega)}
    ,
    \quad 
    u \in L^p(\Omega)
    .
\end{align*}
We have $1 \leq 2^{|2/p-1|} \leq 2$ for $1 \leq p \leq \infty$.
The upper bound is only attained in the limit cases $p = 1$ and $p = \infty$, and the lower bound is only attained if $p = 2$, where we reproduce the improved estimate for the Hilbert space case.
We thus conclude 
\begin{align*}
    C_{\varnothing,\Omega,p} \leq 2^{|2/p-1|} C_{{\rm PF},\Omega,p}.
\end{align*}
In summary, our notion of Poincar\'e--Friedrichs constant is equivalent to the notion of Poincar\'e constant in the literature, up to a numerical factor that only depends on $1 \leq p \leq \infty$.
\\

There is yet another characterization of the Poincar\'e--Friedrichs inequality that we like to point out. 
We define the potential 
\begin{align*}
    \Phi( \bff ) := \argmin\limits_{ \substack{ u \in W^{1,p}(\Omega) \\ \nabla u = \bff } } \| u \|_{L^{p}(\Omega)},
    \quad 
    \bff \in \nabla W^{1,p}(\Omega)
    .
\end{align*}
Clearly, 
\begin{align*}
    \| \Phi(\bff) \|_{L^p(\Omega)} \leq C_{{\rm PF},\Omega,p} \| \bff \|_{L^p(\Omega)},
\end{align*}
and this inequality is sharp by definition of $\Phi$ provided that $C_{{\rm PF},\Omega,p}$ is smallest possible constant in \eqref{math:inequalitypf:gradient}. 
However, $\Phi : \nabla W^{1,p}(\Omega) \rightarrow L^p(\Omega)$ is generally a nonlinear operator for $p \neq 2$. 
Consequently, any \textit{linear} potential operator $\overline\Phi : \nabla W^{1,p}(\Omega) \rightarrow L^p(\Omega)$
satisfying $\nabla \overline\Phi( \bff ) = \bff$ for any $\bff \in \nabla W^{1,p}(\Omega)$ must have an operator norm that obeys the lower bound 
\begin{align*}
    C_{{\rm PF},\Omega,p} 
    \leq 
    \argmin\limits_{ u \in \nabla W^{1,p}(\Omega) \setminus \bbR } 
    \frac{ \| \overline\Phi(\nabla u) \|_{L^p(\Omega)} }{ \| \nabla u \|_{L^p(\Omega)} }
    .
\end{align*}
A natural choice is the linear operator $\Phi_{\varnothing} : \nabla W^{1,p}(\Omega) \rightarrow L^p(\Omega)$ that satisfies 
\begin{align*}
    \Phi_{\varnothing}( \nabla u ) = u - u_{\Omega},
    \quad 
    u \in W^{1,p}(\Omega)
    .
\end{align*}
Its operator norm is just the smallest Poincar\'e constant $C_{\varnothing,\Omega,p}$.
\\

In the light of these musings, we settle for the objective of constructing linear potential operators for the gradient and later for the curl and divergence operators.
This task is more specific than just proving an upper bound for the Poincar\'e--Friedrichs constant. 
Any Poincar\'e--Friedrichs inequality~\eqref{math:inequalitypf:gradient} is comparable to an upper bound for the generalized (possibly nonlinear) inverse of the gradient operator $\nabla : W^{1,p}(\Omega) \rightarrow L^{p}(\Omega)^{n}$. 
The potential reconstruction is the natural approach towards the topic when addressing the curl and divergence operators.

\begin{remark}
    Let us further remark why the above notion of Poincar\'e--Friedrichs inequality~\eqref{math:inequalitypf:gradient} suits our discussion better than the Poincar\'e inequality. 
    We want to generalize the discussion to the curl and divergence operators. 
    The kernel of the gradient is the one-dimensional space of constant functions, and is thus complemented in the Lebesgue spaces with a canonical choice of projection. 
    By contrast, the curl and divergence operators have infinite-dimensional kernels. 
    Hence, it is not clear a priori whether these kernels are complemented subspaces and admit a projection onto them, not to mention a canonical projection. 
    % The kernel of those operators is infinite-dimensional, whereas the gradient only has a one-dimensional kernel. 
    % When discussing Banach spaces, there is no canonical orthogonal projection on the kernel, unlike in the Hilbert space case, 
    % which is why the Poincar\'e--Friedrichs inequality seems to be the more natural approach.
\end{remark}

@@

\subsection{Poincar\'e--Friedrichs inequalities over convex domains}
% See the collection at Lemma 3.24 : https://hal.science/hal-03226049/file/volI_hal.pdf

We collect examples for Poincar\'e and Poincar\'e--Friedrichs inequalities for the important special case of convex domains. 
see also~\cite[Lemma~3.24]{ern2021finite} for a short overview. 
Over convex bounded domains, we have the Poincar\'e inequalities~\cite{bebendorf2003note,acosta2004optimal} 
\begin{align}\label{math:acostaduran}
    \| u - u_{\Omega} \|_{L^{1}(\Omega)}
    \leq 
    \frac{\diam(\Omega)}{2}
    \| \nabla u \|_{L^{1}(\Omega)}
    ,
    \quad 
    u \in W^{1,1}(\Omega)
    ,
    \\
    \| u - u_{\Omega} \|_{L^{2}(\Omega)} \label{math:bebendorf}
    \leq 
    \frac{\diam(\Omega)}{\pi}
    \| \nabla u \|_{L^{2}(\Omega)}
    ,
    \quad 
    u \in W^{1,2}(\Omega)
    .
\end{align}
These two estimates are the best possible Poincar\'e inequalities in the cases $p=1$ and $p=2$, respectively, in terms of the diameter alone. 
Upper bounds for the Poincar\'e constant over convex domains with $1 < p < \infty$ are known in the literature~\cite{chua2006estimates}:
\begin{align}\label{math:chuawheeden}
    \| u - u_{\Omega} \|_{L^{p}(\Omega)}
    \leq 
    C_{{\rm CW},p}
    \diam(\Omega)
    \| \nabla u \|_{L^{p}(\Omega)}
    ,
    \quad 
    u \in W^{1,p}(\Omega)
    ,
\end{align}
where 
\begin{align}
    C_{{\rm CW},p} 
    := 
    \sup\limits_{ v \in C^\infty([0,1]) \setminus \bbR } 
    \frac{ 
        \| v - v_{[0,1]} \|_{L^{p}([0,1])} 
    }{ 
        \| \nabla v \|_{L^{p}([0,1])} 
    }
    \leq 
    \sqrt[p]{p} 2^{1-\frac 1 p}
    =
    2
    \left( \frac p 2 \right)^{\frac 1 p}
    .
\end{align}
These Poincar\'e inequalities imply Poincar\'e--Friedrichs inequalities.

We know optimal Poincar\'e--Friedrichs constants over convex domains~\cite{ferone2012remark,esposito2013poincare}: 
when $1 < p < \infty$, one can show that 
\begin{align}\label{math:esposito:pf}
    \min\limits_{ c \in \bbR }
    \| u - c \|_{L^{p}(\Omega)}
    \leq 
    C_{\rm{EFNT},p}
    \diam(\Omega)
    \| \nabla u \|_{L^{p}(\Omega)}
    ,
    \quad 
    u \in W^{1,p}(\Omega)
    ,
\end{align}
where $C_{\rm{EFNT},p}$ is the best possible constant that only depends on $p$ and equals 
\begin{align}
    C_{\rm{EFNT},p}
    :=
    \frac{ p \sin(\pi/p) }{ 2\pi \sqrt[p]{p-1} }
    .
\end{align}
% TODO: double check what is known in the literature 
Note that the last inequalities imply, again when $1 \leq p < \infty$, the Poincar\'e inequalities
\begin{align}\label{math:esposito:poincare}
    \| u - u_{\Omega} \|_{L^{p}(\Omega)}
    \leq 
    2^{|1-\frac 2 p|}C_{\rm{EFNT},p}
    \diam(\Omega)
    \| \nabla u \|_{L^{p}(\Omega)}
    ,
    \quad 
    u \in W^{1,p}(\Omega)
    .
\end{align}
Finally, since convex domains are Lipschitz domains: 
\begin{align}\label{math:lipschitzestimate}
    \min\limits_{ c \in \bbR }
    \| u - c \|_{L^{\infty}(\Omega)}
    \leq 
    \diam(\Omega)
    \| \nabla u \|_{L^{\infty}(\Omega)}
    ,
    \quad 
    u \in W^{1,\infty}(\Omega)
    .
\end{align}

\begin{remark}
    The main feature in proving these results is that the best constant in the Poincar\'e inequality is already the constant in the $1$-dimensional case. 
    Let us compare the different estimates for the Poincar\'e inequality.
    In the case $2 \leq p$, 
    \begin{align*}
        2^{1-\frac 2 p}
        C_{\rm{EFNT},p}
        =
        \frac{ 2^{1-\frac 2 p} }{2}
        \frac{ \sin(\pi/p) }{ \pi/p }
        \frac{ 1 }{ \sqrt[p]{p-1} }
        =
        4^{-\frac 1 p}
        \leq
        C_{{\rm CW},p} 
        .
    \end{align*}
    In the case $2 \geq p$, 
    \begin{align*}
        2^{\frac 2 p - 1}
        C_{\rm{EFNT},p}
        =
        \frac{ 2^{\frac 2 p - 1} }{2}
        \frac{ \sin(\pi/p) }{ \pi/p }
        \frac{ 1 }{ \sqrt[p]{p-1} }
        \leq
        \frac{ 2^{\frac 2 p-1} }{2}
        =
        2^{\frac 2 p - 2}
        =
        4^{\frac 1 p - 1}
        \leq
        C_{{\rm CW},p} 
%         % 
%         .
%         4^{\frac 1 p}
%         \frac{ p \sin(\pi/p) }{ 2\pi \sqrt[p]{p-1} \cdot 2 }
%         =
%         4^{\frac 1 p}
%         \frac{ p \sin(\pi/p) }{ 4\pi \sqrt[p]{p-1} }
%         \leq 
%         \frac 1 2
        .
    \end{align*}
    It follows that \eqref{math:esposito:poincare} is tighter estimate than \eqref{math:chuawheeden} for $1 < p < \infty$.
    In the limit as $p$ goes to $1$, \eqref{math:esposito:poincare} recovers \eqref{math:acostaduran}. 
    In the limit as $p$ goes to infinity, \eqref{math:esposito:poincare} implies the Poincar\'e inequality 
    \begin{align*}
        \| u - u_{\Omega} \|_{L^{\infty}(\Omega)}
        \leq 
        \diam(\Omega)
        \| \nabla u \|_{L^{\infty}(\Omega)}
        ,
        \quad
        u \in W^{1,\infty}(\Omega)
        %
    \end{align*}
    because the $L^{\infty}$-norm is the limit of the $L^{p}$-norm. 
    Of course, this Poincar\'e inequality implies a Poincar\'e--Friedrichs inequality with the same constant. 
    Notably, it is much less clear whether \eqref{math:esposito:pf} implies a limit case Poincar\'e--Friedrichs inequality for Lipschitz functions. 
%     Taking the \emph{informal} limit also suggest the inequality 
%     \begin{align*}
%         \min\limits_{ c \in \bbR }
%         \| u - c \|_{L^{\infty}(\Omega)}
%         \leq 
%         \frac{d}{2}
%         \| \nabla u \|_{L^{\infty}(\Omega)}
%         ,
%         \quad
%         u \in W^{1,\infty}(\Omega)
%         ,
%     \end{align*}
%     however, we generally cannot exchange the minimum and the limit, so that is strictly informal. 
\end{remark}
